---
title: "Covariate_power"
author: "UofI Metaketa"
date: "1/22/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# PA Feedback

A final concern raised by the review committee concerns statistical power. You note that you can detect a neighborhood effect of .5 standard deviations; this is ordinarily considered a large effect by policy standards. One suggestion to improve power is to obtain prognostic covariates – presumably, the strongest of these would be prior administrative data. That again returns us to the question of whether you have established a working relationship with the Zomba City Council such that you could obtain the necessary covariates to effectively block the randomization.

## Covariates
 1. Income
 2. Health

Here we focus only on two treatments, presuming a test of the relationship between treatment assignment to the common metaketa arm and an alternative arm.


```{r setup }
rm(list=ls())
#install.packages("randomizr")
library(randomizr)    # randomizr package for complete random assignment
library(sandwich)
library(lmtest)
```

Build up the power analysis function piece by piece
```{r step1}
## Base code
N <- 45
numblocks <- 3
blocks <- gl(3,N/3)

## For now making the conservative assumption that covariates do not systematically vary by block
health <- sample(x=1:3, size=N, replace=TRUE)
income <- sample(x=1:5, size=N, replace=TRUE)
sdhealth <- sd(health)
sdincome <- sd(income)

## Also assuming that block membership does not systematically determine control outcome
set.seed(12345)
y0 <- health*(sdhealth) + income*(sdincome) + rnorm(n=N,mean=100,sd=2)
y0S <- y0/sd(y0)
summary(lm(y0S~health+income+blocks)) ## prognostic with R^2 rounghly .5

tau <- .5
y1 <- y0S+tau
Z <- block_ra(block_var=blocks)
Y <- Z*y1 + (1-Z)*y0S
ate1 <- coef(lm(Y~Z+blocks))["Z"]
ate1

healthMD <- health - mean(health)
incomeMD <- income - mean(income)
healthMDb <- health - ave(health,blocks)
tapply(healthMDb,blocks,mean)
incomeMDb <- income - ave(income,blocks)
Yb <- Y - ave(Y,blocks)
Zb <- Z - ave(Z,blocks)

## Now doing covariance adjustment following the Gren Lab SOP (plus our own figuring)
## Need to eventually show that this is unbiased
lm1 <- lm(Yb~Zb*(healthMDb+incomeMDb),singular.ok=FALSE)
results1<-coeftest(lm1, vcov = vcovHC(lm1, type = "HC2"))
results1

### And working on getting standardized effect estimates
lm1a <- lm(scale(Yb)~scale(Zb)*(scale(healthMDb)+scale(incomeMDb)),singular.ok=FALSE)
coeftest(lm1a, vcov = vcovHC(lm1a, type = "HC2"))

## Checking this by hand:
## What is this on sd scale?
## https://stackoverflow.com/questions/24305271/extracting-standardized-coefficients-from-lm-in-r
coef(lm1)["Zb"] * sd(Zb)/sd(Yb)

## Now, test H0: tau=0 when tau grows away from zero.
##Ys <- scale(Y)
##Zs <- scale(Z)
##
##powfn <- function(obsZ,y0,tau,cov1,cov2){
##  # y0 is assumed y0/sd(y0)
##  # covs are assumed mean centered
##  y1 <- y0S+tau
##  Z <- sample(obsZ) ## complete_ra(N=N)
##  Y <- Z*y1 + (1-Z)*y0S
##  Ys <- scale(Y)
##  Zs <- scale(Z)
##  thelm <- lm(Ys~Zs*(cov1+cov2),singular.ok=FALSE)
##  results <- coeftest(thelm, vcov = vcovHC(thelm, type = "HC2"))
##  ## sd(Z) is always the same (here .5)
##  ## return(c(tau=tau,tauS=tau*sd(Zs)/sd(Ys), p=results["Zs",4]))
##  ##return(c(tau=tau, p=results["Zs",4]))
##  return(p=results["Zs",4])
##}
##
#### test function
##powfn(obsZ=Z,y0=y0S,tau=.5,cov1=healthMD,cov2=incomeMD)
```

Now combine all of that work into one big function (eventually delete the preceding).

```{r}
powNfn<- function(N,nsims,thetau,nblocks){
  set.seed(12345)
  blocks <- gl(nblocks,N/nblocks)
  stopifnot(length(blocks)==N)
  cov1 <- sample(x=1:3, size=N, replace=TRUE)
  cov2 <- sample(x=1:5, size=N, replace=TRUE)
  sdcov1 <- sd(cov1)
  sdcov2 <- sd(cov2)
  y0 <- cov1*(sdcov1) + cov2*(sdcov2) + rnorm(n=N,mean=100,sd=2)
  y0S <- y0/sd(y0)
  theZ <- block_ra(block_var=blocks) ## complete_ra(N=N)
  ## cov1MD <- cov1 - mean(cov1)
  ## cov2MD <- cov2 - mean(cov2)
  cov1MDb <- cov1 - ave(cov1,blocks)
  cov2MDb <- cov2 - ave(cov2,blocks)
  y1 <- y0S+tau

  powfn2 <- function(obsZ,y1,y0,tau,cov1,cov2,blocks){
    # y0 is assumed y0/sd(y0)
    # covs are assumed mean centered
    ## Z <- sample(obsZ) ## complete_ra(N=N)
    Z <- unsplit(lapply(split(obsZ,blocks),sample),blocks)
    Y <- Z*y1 + (1-Z)*y0S
    Yb <- Y - ave(Y,blocks)
    Ys <- scale(Y)
    Zb <- Z - ave(Z,blocks)
    Zs <- scale(Z)
    thelm <- lm(Ys~Zs*(cov1+cov2),singular.ok=FALSE)
    results <- coeftest(thelm, vcov = vcovHC(thelm, type = "HC2"))
    ## sd(Z) is always the same (here .5)
    ## return(c(tau=tau,tauS=tau*sd(Zs)/sd(Ys), p=results["Zs",4]))
    ##return(c(tau=tau, p=results["Zs",4]))
    return(p=results["Zs",4])
  }

  theps<-replicate(nsims,powfn2(obsZ=theZ,y1=y1,y0=y0S,tau=thetau,cov1=cov1MDb,cov2=cov2MDb,blocks=blocks))
  stopifnot(any(!is.na(theps)))
  return(mean(theps < .05))

  return(theps)
}

powNfn(N=100,nsims=10,thetau=.5,nblocks=3,cov1=health,cov2=income)


possibleTaus <- seq(0,1,.1)
possibleNs <- c(50,120,220)

parms <- expand.grid(tau=possibleTaus,n=possibleNs)

parms$pow <- apply(parms,1,function(x){ message(paste(parms,collapse=" ")); powNfn(N=x[[2]],nsims=1000,thetau=x[[1]]) })

save(parms,file="parms.rda")
```

