---
title: "Covariate_power"
author: "UofI Metaketa"
date: "1/22/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# PA Feedback

A final concern raised by the review committee concerns statistical power. You note that you can detect a neighborhood effect of .5 standard deviations; this is ordinarily considered a large effect by policy standards. One suggestion to improve power is to obtain prognostic covariates – presumably, the strongest of these would be prior administrative data. That again returns us to the question of whether you have established a working relationship with the Zomba City Council such that you could obtain the necessary covariates to effectively block the randomization.

## Covariates
 1. Income
 2. Health


# Multiple Treatments & Covariates

Mult treatments + covariates.

```{r multiple treatments}
rm(list=ls())
#install.packages("randomizr")
library(randomizr)    # randomizr package for complete random assignment
library(sandwich)
library(lmtest)

## Base code
N <- 50

health <- sample(x=1:3, size=N, replace=TRUE)
income <- sample(x=1:5, size=N, replace=TRUE)
sdhealth <- sd(health)
sdincome <- sd(income)

set.seed(12345)
y0 <- health*(sdhealth) + income*(sdincome) + rnorm(n=N,mean=100,sd=2)
y0S <- y0/sd(y0)
summary(lm(y0S~health+income)) ## prognostic with R^2 rounghly .5

tau <- .5
y1 <- y0S+tau
Z <- complete_ra(N=N)
Y <- Z*y1 + (1-Z)*y0S

ate1 <- coef(lm(Y~Z))["Z"]
ate1

healthMD <- health - mean(health)
incomeMD <- income - mean(income)
lm1 <- lm(Y~Z*(healthMD+incomeMD),singular.ok=FALSE)
results1<-coeftest(lm1, vcov = vcovHC(lm1, type = "HC2"))


lm1a <- lm(scale(Y)~scale(Z)*(scale(healthMD)+scale(incomeMD)),singular.ok=FALSE)
coeftest(lm1a, vcov = vcovHC(lm1a, type = "HC2"))
## What is this on sd scale?
## https://stackoverflow.com/questions/24305271/extracting-standardized-coefficients-from-lm-in-r
coef(lm1)["Z"] * sd(Z)/sd(Y)

## Now, test H0: tau=0 when tau grows away from zero.
Ys <- scale(Y)
Zs <- scale(Z)


powfn <- function(obsZ,y0,tau,cov1,cov2){
  # y0 is assumed y0/sd(y0)
  # covs are assumed mean centered
  y1 <- y0S+tau
  Z <- sample(obsZ) ## complete_ra(N=N)
  Y <- Z*y1 + (1-Z)*y0S
  Ys <- scale(Y)
  Zs <- scale(Z)
  thelm <- lm(Ys~Zs*(cov1+cov2),singular.ok=FALSE)
  results <- coeftest(thelm, vcov = vcovHC(thelm, type = "HC2"))
  ## sd(Z) is always the same (here .5)
  ## return(c(tau=tau,tauS=tau*sd(Zs)/sd(Ys), p=results["Zs",4]))
  ##return(c(tau=tau, p=results["Zs",4]))
  return(p=results["Zs",4])
}

## test function
powfn(obsZ=Z,y0=y0S,tau=.5,cov1=healthMD,cov2=incomeMD)

powNfn<- function(N,nsims,thetau){
  health <- sample(x=1:3, size=N, replace=TRUE)
  income <- sample(x=1:5, size=N, replace=TRUE)
  sdhealth <- sd(health)
  sdincome <- sd(income)
  set.seed(12345)
  y0 <- health*(sdhealth) + income*(sdincome) + rnorm(n=N,mean=100,sd=2)
  y0S <- y0/sd(y0)
  theZ <- complete_ra(N=N)
  healthMD <- health - mean(health)
  incomeMD <- income - mean(income)

  powfn2 <- function(obsZ,y0,tau,cov1,cov2){
    # y0 is assumed y0/sd(y0)
    # covs are assumed mean centered
    y1 <- y0S+tau
    Z <- sample(obsZ) ## complete_ra(N=N)
    Y <- Z*y1 + (1-Z)*y0S
    Ys <- scale(Y)
    Zs <- scale(Z)
    thelm <- lm(Ys~Zs*(cov1+cov2),singular.ok=FALSE)
    results <- coeftest(thelm, vcov = vcovHC(thelm, type = "HC2"))
    ## sd(Z) is always the same (here .5)
    ## return(c(tau=tau,tauS=tau*sd(Zs)/sd(Ys), p=results["Zs",4]))
    ##return(c(tau=tau, p=results["Zs",4]))
    return(p=results["Zs",4])
  }

  theps<-replicate(nsims,powfn2(obsZ=theZ,y0=y0S,tau=thetau,cov1=healthMD,cov2=incomeMD))
  stopifnot(any(!is.na(theps)))
  return(mean(theps < .05))

  return(theps)
}

powNfn(N=100,nsims=10,thetau=.5)


possibleTaus <- seq(0,1,.1)
possibleNs <- c(50,120,220)

parms <- expand.grid(tau=possibleTaus,n=possibleNs)

parms$pow <- apply(parms,1,function(x){ message(paste(parms,collapse=" ")); powNfn(N=x[[2]],nsims=1000,thetau=x[[1]]) })

save(parms,file="parms.rda")
```

